\section*{2009}
\vspace{-.5cm}
\hrulefill \\\\
%\addcontentsline{toc}{section}{\currentname}
\textbf{1.(a) (12 marks)} In a two-tier competitive examination, only candidates with a minimum score of $x_0$ in the preliminary examination are allowed to sit for the main examination. The score distribution in the preliminary examination can be assumed to be normal with mean $\mu$ and s.d. $\sigma$. \vspace{.05cm}

Derive the expressions for mean and variance of scores in the preliminary examination of candidates sitting for the main examination. 
\\\\
\textbf{1.(b) (12 marks)} 2 sets of $n$ cards, both numbered from $1$ to $n$, are randomly matched. Find \(E(X)\) and \(V(X)\) where \(X\) is the number of matches realised.
\\\\
\textbf{2.(a) (20 marks)} $X_1$ and $X_2$ are i.i.d. random variables with common probability mass function
\begin{equation*}
\begin{aligned}
    f(x) = &q^x p, \qquad &&x = 0,1,2, .....  ,\\
           &        &&0 < p < 1, q = 1 - p,\\
         = &0,      &&otherwise.
\end{aligned}
\end{equation*}
Obtain the probability distribution of \( Y = max(X_1, X_2)\). Also derive $E(Y)$.
\\\\
\textbf{4.(a) (20 marks)} The daily demand (in kg) of fish of a certain variety upto 12 noon with a retailer follows exponential law with mean 25. He makes a profit of Rs.20 on an average for each kg he is able to sell by 12 noon, and incurs a loss of Rs.10 on an average for each kg that remains unsold after 12 noon.\vspace{.05cm}

What is the optimum daily stock he should procure.
\\\\
\textbf{4.(b) (20 marks)} For the pareto distribution with distribution function
\begin{equation*}
    f(x) = \begin{cases}
    1 - (\theta / x)^\alpha \qquad&, x \geq \theta\\
    0 &, otherwise.
    \end{cases}
\end{equation*}
Derive the p.d.f. Also obtain its harmonic mean and variance.\\


