\section*{2010}
%\addcontentsline{toc}{section}{\currentname}
\vspace{-.5cm}
\hrulefill \\\\
\textbf{1.(a) (10 marks).} 5\% residents of a locality are earning more than Rs 10 lakhs per annum. The percentage of single earner in the family among those earning more than Rs 10 lakhs per annum is 80\%, whereas percentage of single earner among others (earning less than Rs 10 lakhs per annum) is 50\%. A randomly selected resident of that locality was found to be single earner. What is the probability that he earns more than Rs 10 lakhs per annum? 
\\\\
\textbf{1.(b) (10 marks)} \(T\) is a triangle in $(X,Y)$ plane with vertices $(0,0)$, $(0,\sqrt{2})$ and $(\sqrt{2}, \sqrt{2})$. Let $F(x,y)$ denote the area of intersection of T and the set $S =\{(a,b): a\leq x,b\leq y\}$. Does $F$ define a cumulative distribution function? If yes, find the joint density function.
\\\\
\ques{1}{c}{10} Two players $A$ and $B$ decide to play a series of at most 7 games. A player who wins 4 games wins the series. IF both the players have equal chance of wining a game and $A$ wins the first two games, what is the probability that A will win the series? 
\\\\
\ques{2}{a}{20} Cumulative distribution function (c.d.f) of a random variable X is 
\begin{equation*}
    F(x) = \begin{cases}
             0,\quad &if\quad x<0\\
             \frac{1}{4},\quad &if\quad 0\leq x<0\\
             \frac{x}{3},\quad &if\quad 1\leq x<2\\
             1,\quad &if\quad x\geq 2
    \end{cases}
\end{equation*}
Find $\alpha$ such that we can write $F(x) = \alpha F_c(x) + (1 -\alpha)F_d(x)$, where $F_c(x)$ and $F_d(x)$ are c.d.f. of continuous and discrete random variables respectively. $F_c(x)$ and $F_d(x)$.



