\section*{2009}
\vspace{-.5cm}
\hrulefill \smallskip\\
\ques{5}{a}{12} A committee of 5 is to be formed from amongst 86 officers, numbered serially from 1 to 86. Draw a circular systematic sample for this purpose, giving the procedure you have followed in detail. You may use the following set of random numbers : \\
\begin{tabular}{cccccccc}
1349 & 0417 & 9311 & 9787 & 1284 & 0769 & 8422 & 1077.
\end{tabular} 
\myline
\ques{5}{c}{12} Locate 5 points at random on a rectangular paddy field measuring 30m x 20m for conducting crop-cutting experiment. The points are to be located using rectangular co-ordinates to nearest decimeters with one corner of the plot as origin. The following set of random numbers may be used:
\begin{tabular}{cccccccc}
2292 & 2933 & 6125 & 2464 & 1038 & 3163 & 3569 & 7155\\
2029 & 2538 & 7080 & 3027 & 6215 & 3125 & 5856 & 9543.
\end{tabular} 
\myline
\ques{5}{d}{12} For pps sampling without replacement, give the Horwitz-Thompson estimator of the population total. Show that it is unbiased. Also obtain its variance.
\myline
\ques{7}{b}{20} A rural block in a district was divided into 3 strata. The following table gives some relevant information:
\begin{center}
    \begin{tabular}{|p{2cm}|p{2.2cm}|p{2.3cm}|p{2cm}|}
        \hline
         Stratum No. (h) & Number of Villges ($N_h$) & Area $X_h$ under wheat (hectares) & $S.D.S_{X_h}$ (hectares) \\ \hline
         1 & 20 & 620 & 100.5 \\
         2 & 15 & 580 & 90.8 \\ 
         3 & 25 & 825 & 140.7 \\ \hline
    \end{tabular}
\end{center} A sample of 10 villages is to be drawn. How many villages should be selected from different strata using (i) proportional allocation, (ii) optimum allocation, and (iii) allocation proportional to area under crop?
\myline
\ques{8}{a}{20} You are to estimate the proportion of school going students (classes V to XII) using mobile phones in a metropolitan city. Suggest a two-stage sampling scheme. Using suitable notation, give the estimator to be used, and check if the estimator is unbiased.
