\newpage
\section*{2010}
\vspace{-.5cm}
\hrulefill \smallskip\\
\ques{5}{a}{12} Consider a population of 3 units $U_1$, $U_2$ and $U_3$. The size measures of $U_1$, $U_2$ and $U_3$ are respectively $X_1 = 6$, $X_2=4$, and $X_3 = 2$. A sample of two units is drawn from the population without replacement such that the first unit is selected with probability proportional to their sizes and second unit is selected with probability proportional to the sizes of the remaining units. If $\pi_i$ denotes the probability of inclusion of $U_i$ in the sample, show that
\[ \pi_1 = \frac{51}{60}, \enskip \pi_2 = \frac{44}{60}, \enskip\pi_3 = \frac{25}{60}.\]
\ques{1}{e}{12} In a population consisting of three units $U_1$, $U_2$ and $U_3$, the observations are $Y_1$, $Y_2$ and $Y_3$ respectively. A sample of size 2 is drawn from the population by using simple random sampling without replacement. Define an estimator T as follows :
\[ T = \begin{cases} \frac{1}{2}Y_1 + \frac{1}{2}Y_2, \enskip \text{if } U_1 \text{ and } U_2 \text{ selected} \\
\frac{1}{2}Y_1 + \frac{1}{3}Y_3, \enskip \text{if } U_1 \text{ and } U_3 \text{ selected} \\
\frac{1}{2}Y_2 + \frac{1}{3}Y_3, \enskip \text{if } U_2 \text{ and } U_3 \text{ selected}
\end{cases}
\] Show that $T$ is unbiased estimator of population mean and has smaller variance than sample mean if $Y_3(3Y_2 - 3Y_1 - Y_3) > 0$.
\myline
\ques{6}{c}{20} From a population of $N$ units, a random sample of 3 unis is drawn by simple random sampling with replacement. Find the expectation of distinct units in the sample. Define $T$ as mean of distinct units in the sample. Show that $T$ is unbiased estimate of population mean.
\myline
\ques{7}{b}{20} The variable $X$ under study has rectangular distribution in the interval $(a, a+d)$. The interval is divided into $k$ equal sub intervals which form k strata of equal sizes. From each stratum, a simple random sample of $n/k$ units is drawn. Let $V_1$ and $V_2$ be the variances of estimator of population mean based on stratified and unstratified samples of size $n$ respectively. Prove that $V_1/V_2 = k^{-2}$.
\myline
\ques{8}{a}{20} Define ratio estimator for estimating the population total and derive expression for the standard error of the estimator.