\section*{2013}
\vspace{-.5cm}
\hrulefill \smallskip\\
\ques{5}{c}{10} Show that systematic sample is better than SRSWOR if $-\dfrac{1}{n -1} <\rho_w <-\dfrac{1}{N-1}$ where $n$ and $N$ are sample and population sizes respectively and $\rho_w$ is the intraclass correlation between pairs of units that are in the same systematic sample.
\myline
\ques{6}{c}{15} Suppose it is required to estimate the average value of output of a group of 5000 factories in a region so that the sample estimate lies within 10\% of the true value with confidence coefficient of 95\%. The population coefficient of variation is known to be  60\%. Determine the equation to calculate minimum sample size.
\myline
\ques{7}{c}{15} A rural block in a district was divided into 3 strata. The following table gives the number of villages $(N_h)$ and standard deviation $(S_h)$ of area under wheat crop for different strata:
\begin{center}
 \begin{tabular}{|*{3}{c|}} \hline
    Stratum No & No of Villages($N_h$) & Standard deviation($S_h$) \\ \hline
    1 & 15 & 100 \\ \hline
    2 & 10 & 50 \\ \hline
    3 & 25 & 120 \\ \hline
\end{tabular}   
\end{center} A sample of 20 villages is to be drawn. How many villages should be selected under each strata under \begin{tasks}
\task proportional allocation?
\task Neyman allocation? \end{tasks}
\ques{8}{c}{15} Define ratio estimator. Show that it is biased. Suggest a suitable sampling procedure for which the classical ratio estimator of a population ratio is unbiased. For the suggested procedure, propose an unbiased estimator of the variance of ratio estimator.