\section*{2018}
\vspace{-.5cm}
\hrulefill \smallskip\\
\ques{1}{d}{10} A population has seven units consisting of 30, 10 ,25,5,20,6 and 4 fields. Select a PPS sample of units of size 3, using random numbers 72,46 ad 94.
\myline
\ques{6}{c}{15} Compare ratio and regression methods of estimation. Show that ratio method gives a biased estimation of population level. Derive an approximate expression of the bias.
\myline
\ques{7}{c}{15} Two dentists A and B make a survey of the state of the teeth of 200 children in a village. Dr. A selects a simple random sample of 20 children and counts the  number of decayed teeth for each child with the following results
\begin{center}
    \begin{tabular}{|l|*{11}{c|}}\hline
        No. of decayed teeth & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10\\ \hline
        No. of children & 8 & 4 & 2 & 2 & 1 & 1 & 1 & 0 & 0 & 1 & 1\\\hline
    \end{tabular}
\end{center} Dr B using the same dental technique, examines all 200 children, recording that merely those who have no decayed teeth. He finds 60 children with no decayed teeth.\\ Estimate the total number of decayed teeth using
\begin{enumerate}[topsep=0pt, itemsep = -1ex,label=(\roman*)]
    \item Dr. A's results only
    \item Both Dr. A's and Dr. B's results
\end{enumerate} Which estimate do you prefer and why?
\myline
\ques{8}{a}{20} Explain stratified random sampling and give reasons for its preferenece to simple random sampling. \\\\ With two strata, a sampler would have to have $n_1 = n_2$ for administrative convenience instead of using the values given by Neyman allocation. If $V(\bar{y}_{st})$ and $V_{opt}(\bar{y}_{st})$ denote the variances given by $n_1 = n_2$ and Neyman allocations respectively, show that the fractional increase in the variaance is
\[ \dfrac{V(\bar{y}_{st}) - V_{opt}(\bar{y}_{st})}{V_{opt}(\bar{y}_{st})} = \left(\dfrac{r-1}{r+1}\right)^2 \] where $r=n_1/n_2$ as given by Neyman allocation.