\section*{2009}
\vspace{-.5cm}
\hrulefill \smallskip\\
\ques{1}{e}{12} What is a size-$\alpha$ randomized test?\smallskip \\
For testing the null hypothesis that at most 50\% of the 500ml. packs of hair oil of a certain brand are underweight, against the alternative that the percentage is more, it is desired to examine 8 such packs chosen at random. Use the following information to design a randomized test of size 0.05 :
\begin{center}
    \begin{tabular}{|*{10}{c|}}
        \hline
        x & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8  \\\hline
        f(x) & .0039 &.0312 & .1094 & .2188 & .2934 & .2188 & .1094 & .0312 & .0039\\\hline 
    \end{tabular}
\end{center}
where f(x) is the probability mass function of a binomial variate with parameters m=8, p=0.5.
\myline
\ques{1}{f}{12} A group of 20 short distance runners were subjected to a month-long training. Discuss how you would examine if the training was at all effective based on their timings to clear 100 meters before and after the training, clearly mentioning the underlying assumptions, if any, the null and alternative hypothesis, as also the size-$\alpha$ critical region.
\myline
\ques{2}{b}{20} In sampling from N($\mu$, $\sigma^2 $), $\mu$ unknown, derive 100(1 - $\alpha$)\% shortest confidence interval for $\sigma^2$.
\myline
\ques{3}{c}{20} Show that the likelihood ratio test, for the hypothesis that the means of K independent normal populations with common variance are identical, boils down to a F-test.
\myline
\ques{4}{c}{20} Define (i) most powerful test, and (ii) unbiased test of a simple hypothesis against a simple alternative.
\par Show that a most powerful test is necessarily unbiased.